---
title: "6-Gene Ontology enrichment"
bibliography: bib.bib
link-citations: yes
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output: 
  pdf_document:
    latex_engine: xelatex
    pandoc_args: 
      - '../../common.yaml'
      - --listings
    includes: 
      in_header: '../../preamble.tex'
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
    df_print: tibble
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


\newpage

# Description

The DE gene lists used for enrichment analyses were calculated using Seurat function FindMarkers[@SeuratV4] with default arguments except only.pos = TRUE in order to output only positively regulated genes and logfc.threshold of 0.2. KEGG and Gene Ontology enrichment analyses were made using enrichKEGG and enrichGO functions from package clusterProfiler package[@clusterProfiler]⁠ with default arguments. 

# Load data and packages 

```{r eval=FALSE}
library(Seurat)
library(ggplot2)
library(dplyr)
results <- readRDS(file = "../3-Merge and cell typing/so.merged_clusters.seuratObject.Rds")
```

```{r, remove after render-load data, include=FALSE}
library(Seurat)
library(ggplot2)
library(dplyr)
results <- readRDS(file = "../../../Human_Monocytes/4 - cell typing and filtering/so.merged_clusters.seuratObject.Rds")
```

# DE genes and GO enrichment

## Cluster 1 DE analysis and GO enrichment

> The Cluster 1 mentioned in manuscript refers to cluster 0 in the codes following. The same for cluster 2-4 in manuscript, they are cluster 1-3 in the codes. 

Calculate the DE genes. 

```{r}
de_ident0 <- FindMarkers(results, ident.1 = "0", only.pos = TRUE, logfc.threshold = 0.2)
head(de_ident0)
```

Save data

```{r}
write.csv(de_ident0, file = "./Results_lists/DE_cluster1.csv")
```

KEGG/GO enrichment: 

```{r echo=TRUE}
library(clusterProfiler)
source("../../R/entrez2symbol.R")
source("../../R/replaceEntrezID.R")

de_entrez.ident0 <- bitr( geneID = rownames(de_ident0), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db", drop = TRUE ) $ ENTREZID
result.enrichKEGG.de_ident0 <- enrichKEGG(de_entrez.ident0, organism = "hsa", keyType = "ncbi-geneid")
result.enrichKEGG.de_ident0 <- replaceEntrezID(result.enrichKEGG.de_ident0)
```


```{r}
result.enrichKEGG.de_ident0@result
```

```{r}
write.csv(result.enrichKEGG.de_ident0, file = "./Results_lists/enrichKEGG_DE_cluster1.csv")
```


```{r}
result.enrichGO.de_ident0 <- enrichGO(gene = de_entrez.ident0, OrgDb = "org.Hs.eg.db", ont = "BP")
result.enrichGO.de_ident0 <- replaceEntrezID(result.enrichGO.de_ident0)
result.enrichGO.de_ident0@result
```

```{r}
write.csv(result.enrichGO.de_ident0, file = "./Results_lists/enrichGOBP_DE_cluster1.csv")
```

## Cluster 2 DE analysis and GO enrichment

> The Cluster 2 mentioned in manuscript refers to cluster 1 in the codes following. 

Calculate the DE genes. 

```{r}
de_ident1 <- FindMarkers(results, ident.1 = "1", only.pos = TRUE, logfc.threshold = 0.2)
head(de_ident1)
```

Save data

```{r}
write.csv(de_ident1, file = "./Results_lists/DE_cluster2.csv")
```

KEGG/GO enrichment: 

```{r echo=TRUE}
de_entrez.ident1 <- bitr( geneID = rownames(de_ident1), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db", drop = TRUE ) $ ENTREZID
result.enrichKEGG.de_ident1 <- enrichKEGG(de_entrez.ident1, organism = "hsa", keyType = "ncbi-geneid")
result.enrichKEGG.de_ident1 <- replaceEntrezID(result.enrichKEGG.de_ident1)
```


```{r}
result.enrichKEGG.de_ident1@result
```

```{r}
write.csv(result.enrichKEGG.de_ident1, file = "./Results_lists/enrichKEGG_DE_cluster2.csv")
```


```{r}
result.enrichGO.de_ident1 <- enrichGO(gene = de_entrez.ident1, OrgDb = "org.Hs.eg.db", ont = "BP")
result.enrichGO.de_ident1 <- replaceEntrezID(result.enrichGO.de_ident1)
result.enrichGO.de_ident1@result
```

```{r}
write.csv(result.enrichGO.de_ident1, file = "./Results_lists/enrichGOBP_DE_cluster2.csv")
```


## Cluster 3 DE analysis

> The Cluster 3 mentioned in manuscript refers to cluster 2 in the codes following. 

Calculate the DE genes. 

```{r}
de_ident2 <- FindMarkers(results, ident.1 = "2", only.pos = TRUE, logfc.threshold = 0.2)
nrow(de_ident2)
```

Save data

```{r}
write.csv(de_ident2, file = "./Results_lists/DE_cluster3.csv")
```

## Cluster 4 DE analysis

> The Cluster 4 mentioned in manuscript refers to cluster 3 in the codes following. 

Calculate the DE genes. 

```{r}
de_ident3 <- FindMarkers(results, ident.1 = "3", only.pos = TRUE, logfc.threshold = 0.2)
head(de_ident3)
```

Save data

```{r}
write.csv(de_ident3, file = "./Results_lists/DE_cluster4.csv")
```


## GO/KEGG enrichment and DE analysis in subpopulations of Cluster 3 after re-clustering

Load Cluster 3 object: 

```{r include=FALSE}
results.c2 <- readRDS("../../../Human_Monocytes/9 - Analysis of cluster2 (monocytes-derived)/cluster2_clustered.seuratObject.Rds")
```
```{r eval=FALSE}
results.c2 <- readRDS("../4-Functional characterization of clustered populations/cluster2_clustered.seuratObject.Rds")
```

> REMIND: the subpopulations cluster 1-4 (after reclustering the cluster 3) in the manuscript refer to the cluster 0-3 in the following codes.

Calculate the DE genes. 

DE genes for each subpopulations: 

```{r}
de_subpop0.c2 <- FindMarkers(object = results.c2, ident.1 = "0", only.pos = TRUE, logfc.threshold = 0.2, verbose = FALSE)
de_subpop1.c2 <- FindMarkers(object = results.c2, ident.1 = "1", only.pos = TRUE, logfc.threshold = 0.2, verbose = FALSE)
de_subpop2.c2 <- FindMarkers(object = results.c2, ident.1 = "2", only.pos = TRUE, logfc.threshold = 0.2, verbose = FALSE)
de_subpop3.c2 <- FindMarkers(object = results.c2, ident.1 = "3", only.pos = TRUE, logfc.threshold = 0.2, verbose = FALSE)
```

Save data

```{r}
write.csv(de_subpop0.c2, file = "./Results_lists/DE_subpop1_c3.csv")
write.csv(de_subpop1.c2, file = "./Results_lists/DE_subpop2_c3.csv")
write.csv(de_subpop2.c2, file = "./Results_lists/DE_subpop3_c3.csv")
write.csv(de_subpop3.c2, file = "./Results_lists/DE_subpop4_c3.csv")
```

Since the subpopulations 2 and 3 are very similar. We would like also to have the DE genes in both subpopulations 2 & 3 comparing to subpopulation 1 (we excluded cluster 4 because it's high similarity to DCs). 

```{r}
de_subpop1_2.c2 <- FindMarkers(object = results.c2, ident.1 = c("1", "2"), ident.2 = ("0"), only.pos = TRUE, logfc.threshold = 0.2, verbose = FALSE)
write.csv(de_subpop1_2.c2, file = "./Results_lists/DE_subpop2_and_3_of_c3.csv")
```

KEGG enrichment: 

```{r}
de_entrez.subpop0.c2 <- bitr( geneID = rownames(de_subpop0.c2), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db", drop = TRUE ) $ ENTREZID
de_entrez.subpop1.c2 <- bitr( geneID = rownames(de_subpop1.c2), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db", drop = TRUE ) $ ENTREZID
de_entrez.subpop2.c2 <- bitr( geneID = rownames(de_subpop2.c2), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db", drop = TRUE ) $ ENTREZID
de_entrez.subpop3.c2 <- bitr( geneID = rownames(de_subpop3.c2), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db", drop = TRUE ) $ ENTREZID
de_entrez.subpop1_2.c2 <- bitr( geneID = rownames(de_subpop1_2.c2), fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Hs.eg.db", drop = TRUE ) $ ENTREZID

result.enrichKEGG.de_entrez.subpop0.c2 <- enrichKEGG(de_entrez.subpop0.c2, organism = "hsa", keyType = "ncbi-geneid")
result.enrichKEGG.de_entrez.subpop1.c2 <- enrichKEGG(de_entrez.subpop1.c2, organism = "hsa", keyType = "ncbi-geneid")
result.enrichKEGG.de_entrez.subpop2.c2 <- enrichKEGG(de_entrez.subpop2.c2, organism = "hsa", keyType = "ncbi-geneid")
result.enrichKEGG.de_entrez.subpop3.c2 <- enrichKEGG(de_entrez.subpop3.c2, organism = "hsa", keyType = "ncbi-geneid")
result.enrichKEGG.de_entrez.subpop1_2.c2 <- enrichKEGG(de_entrez.subpop1_2.c2, organism = "hsa", keyType = "ncbi-geneid")

result.enrichGO.de_entrez.subpop0.c2 <- enrichGO(gene = de_entrez.subpop0.c2, OrgDb = "org.Hs.eg.db", ont = "BP")
result.enrichGO.de_entrez.subpop1.c2 <- enrichGO(gene = de_entrez.subpop1.c2, OrgDb = "org.Hs.eg.db", ont = "BP")
result.enrichGO.de_entrez.subpop2.c2 <- enrichGO(gene = de_entrez.subpop2.c2, OrgDb = "org.Hs.eg.db", ont = "BP")
result.enrichGO.de_entrez.subpop3.c2 <- enrichGO(gene = de_entrez.subpop3.c2, OrgDb = "org.Hs.eg.db", ont = "BP")
result.enrichGO.de_entrez.subpop1_2.c2 <- enrichGO(gene = de_entrez.subpop1_2.c2, OrgDb = "org.Hs.eg.db", ont = "BP")

result.enrichKEGG.de_entrez.subpop0.c2 <- replaceEntrezID(result.enrichKEGG.de_entrez.subpop0.c2)
result.enrichKEGG.de_entrez.subpop1.c2 <- replaceEntrezID(result.enrichKEGG.de_entrez.subpop1.c2)
result.enrichKEGG.de_entrez.subpop2.c2 <- replaceEntrezID(result.enrichKEGG.de_entrez.subpop2.c2)
result.enrichKEGG.de_entrez.subpop3.c2 <- replaceEntrezID(result.enrichKEGG.de_entrez.subpop3.c2)
result.enrichKEGG.de_entrez.subpop1_2.c2 <- replaceEntrezID(result.enrichKEGG.de_entrez.subpop1_2.c2)

result.enrichGO.de_entrez.subpop0.c2 <- replaceEntrezID(result.enrichGO.de_entrez.subpop0.c2)
result.enrichGO.de_entrez.subpop1.c2 <- replaceEntrezID(result.enrichGO.de_entrez.subpop1.c2)
result.enrichGO.de_entrez.subpop2.c2 <- replaceEntrezID(result.enrichGO.de_entrez.subpop2.c2)
result.enrichGO.de_entrez.subpop3.c2 <- replaceEntrezID(result.enrichGO.de_entrez.subpop3.c2)
result.enrichGO.de_entrez.subpop1_2.c2 <- replaceEntrezID(result.enrichGO.de_entrez.subpop1_2.c2)
```

```{r}
write.csv(result.enrichKEGG.de_entrez.subpop0.c2, file = "./Results_lists/enrichKEGG_DE_subpop1_c3.csv")
write.csv(result.enrichKEGG.de_entrez.subpop1.c2, file = "./Results_lists/enrichKEGG_DE_subpop2_c3.csv")
write.csv(result.enrichKEGG.de_entrez.subpop2.c2, file = "./Results_lists/enrichKEGG_DE_subpop3_c3.csv")
write.csv(result.enrichKEGG.de_entrez.subpop3.c2, file = "./Results_lists/enrichKEGG_DE_subpop4_c3.csv")
write.csv(result.enrichKEGG.de_entrez.subpop1_2.c2, file = "./Results_lists/enrichKEGG_DE_subpop2_3.c3.csv")

write.csv(result.enrichGO.de_entrez.subpop0.c2, file = "./Results_lists/enrichGOBP_DE_subpop1_c3.csv")
write.csv(result.enrichGO.de_entrez.subpop1.c2, file = "./Results_lists/enrichGOBP_DE_subpop2_c3.csv")
write.csv(result.enrichGO.de_entrez.subpop2.c2, file = "./Results_lists/enrichGOBP_DE_subpop3_c3.csv")
write.csv(result.enrichGO.de_entrez.subpop3.c2, file = "./Results_lists/enrichGOBP_DE_subpop4_c3.csv")
write.csv(result.enrichGO.de_entrez.subpop1_2.c2, file = "./Results_lists/enrichGOBP_DE_subpop2_3_C3.csv")
```

Head 10 lines in GO: 

For subpop 1: 
```{r}
head(result.enrichGO.de_entrez.subpop0.c2@result, 10)
```

For subpop 2: 
```{r}
head(result.enrichGO.de_entrez.subpop1.c2@result, 10)
```

For subpop 3: 
```{r}
head(result.enrichGO.de_entrez.subpop2.c2@result, 10)
```

For subpop 4: 
```{r}
head(result.enrichGO.de_entrez.subpop3.c2@result, 10)
```

For subpop 2,3: 
```{r}
head(result.enrichGO.de_entrez.subpop1_2.c2@result, 10)
```

# Session information 

R sesssion: 

```{r}
sessionInfo()
```

# References
